SciPy 1.0---Fundamental Algorithms for Scientific Computing in Python

### 2019-02-22 scipy-1.0/paper.tex
Change roadmap summary of ndimage
removed unnecessary comma

% PhD Mayo Clinic, started faculty position at BYU in 2001
By the late 1990s, discussions appeared on Matrix-SIG,
expressing a desire for a complete scientific data analysis environment\cite{Travis-Keynote-2010}.
Travis Oliphant, a PhD student at the Mayo Clinic,
released a number of packages\cite{Travis-some-modules,Travis-enhance} 
that built on top of the Numeric array
package, and provided algorithms for signal processing, special
functions, sparse matrices, quadrature, optimization, Fast Fourier
Transforms, and more.  One of these packages, Multipack, was a set of
extension modules that wrapped Fortran and C libraries such as
ODEPACK, QUADPACK, and MINPACK, to solve and minimize nonlinear
equations, integrate differential equations, and fit splines (the
latter implemented by Pearu Peterson).  Robert Kern, then an
undergraduate student (and currently a SciPy core developer), provided
compilation instructions under Windows.
% 2001 PhD Tallinn Technical University
Around the same time, Pearu Peterson, a PhD student from Estonia,
released f2py\cite{Pearu-f2py-mail}, a command line tool for binding Python and Fortran
codes, and wrote modules for linear algebra and interpolation.
% Graduated 1999, Duke
Eric Jones, while a graduate student at Duke, wrote a number of
packages to support his dissertation, including a parallel job
scheduler and genetic optimizer.
% Graduated 1998, PhD from Brown, 97–99 postdoc at Harvard Medical
% School; at Harvard ever since
%
% https://mail.python.org/pipermail/scipy-dev/2001-November/000088.html
%
Gary Strangman, a postdoctoral fellow at Harvard Medical School,
published several descriptive and inferential statistical 
routines\cite{Strangman-modules}.

@@

% separate Figure / Table numbering for SI
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}

@@

\begin{table}[H]
  \centering
  \caption{Optimization methods from \texttt{minimize}, which solves problems of the form $\min_x f(x)$, where $x \in \mathbb{R}^n$ and $f: \mathbb{R}^n \rightarrow \mathbb{R}$ .  The field \textit{version added} specifies the algorithm's first appearance in SciPy. Algorithms with \textit{version added} ``0.6*'' were added in version 0.6 or before.
    The field \textit{wrapper} indicates whether the implementation available in SciPy wraps a function written in a compiled language
    (e.g. C or FORTRAN). The fields \textit{\nth{1}} and \textit{\nth{2} derivatives}
    indicates whether first or second order derivatives are required. When \textit{\nth{2} derivatives} is flagged
    with $\sim$ the algorithm does not requires second-order derivatives from
    the user; it computes an approximation internally and uses it to accelerate method convergence.
    \textit{Iterative Hessian factorization} denotes algorithms that factorize the Hessian in an iterative way,
    which does not require explicit matrix factorization or storage of the Hessian.
    \textit{Local convergence} gives a lower bound on the rate of convergence of the iterations sequence once the
    iterate is sufficiently close to the solution: linear (L), superlinear (S) and quadratic (Q). Convergence rates denoted S$^*$ indicate that the algorithm
    has a superlinear rate for the parameters used in SciPy, but can  achieve a quadratic convergence rate with other parameter choices.
    \textit{Global convergence} is marked for the algorithms with guarantees of convergence to a stationary
    point (i.e. a point $x^*$ for which $\nabla f(x^*) = 0$); this is \emph{not} a guarantee of convergence to a global minimum. The table also indicates which algorithms
    can deal with constraints on the variables. We distinguish between: \textit{bound constraints} (i.e. $x^l \le x \le x^u$),
    \textit{equality constraints} (i.e. $c_{\text{eq}}(x) = 0$) and \textit{inequality constraints} (i.e. $c_{\text{ineq}}(x) \ge 0$).}
  \begin{tabular}{cccccccccccccc}
      & \rotatebox{80}{\texttt{Nelder-Mead}} & \rotatebox{80}{\texttt{Powell}} & \rotatebox{80}{\texttt{COBYLA}} & \rotatebox{80}{\texttt{CG}} & \rotatebox{80}{\texttt{BFGS}}&  \rotatebox{80}{\texttt{L-BFGS-B}} & \rotatebox{80}{\texttt{SLSQP}} & \rotatebox{80}{\texttt{TNC}} & \rotatebox{80}{\texttt{Newton-CG}} & \rotatebox{80}{\texttt{dogleg}} & \rotatebox{80}{\texttt{trust-ncg}} & \rotatebox{80}{\texttt{trust-exact}} & \rotatebox{80}{\texttt{trust-krylov}} \\
    \hline
    Version added &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.6* &  0.9 &  0.6* &  0.6* & 0.13 & 0.13 & 0.19 & 1.0 \\
    \hline
    Wrapper & & & \cmark & & & \cmark & \cmark & \cmark & &  & & & \cmark \\
    \hline
    \nth{1} derivatives &  & & & \cmark  & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
    \hline
    \nth{2} derivatives &  &  &  &  & $\sim$ & $\sim$ & $\sim$ & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark \\
    \hline
    \makecell{Iterative Hessian \\
    factorization} & & & &  & & & & \cmark & \cmark &  & \cmark &  & \cmark \\
    \hline
    Local convergence& & & & L & S &  L & S & S$^*$ & S$^*$ & Q & S$^*$ & Q & S$^*$  \\
    \hline
    Global convergence & & &  &   & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark & \cmark  \\
    \hline
    \makecell{Line-search (LS) or\\ trust-region (TR)} & Neither  & LS &  TR & LS & LS & LS & LS & LS & LS & TR & TR & TR & TR \\
    \hline
    Bound constraints &&&\cmark&&&&\cmark&\cmark&\cmark&&&& \\
    \hline
    Equality constraints &&&&&&&\cmark&&&&& \\
    \hline
    Inequality constraint &&&\cmark&&&&\cmark&&&&& \\
    \hline
    References & \cite{nelder_simplex_1965, wright_direct_1996} & \cite{powell_efficient_1964} &
      \cite{powell_direct_1994, powell_direct_1998, powell_view_2007} &
      \cite{polak_note_1969, nocedal_numerical_2006} & \cite{nocedal_numerical_2006} & \cite{byrd_limited_1995, zhu_algorithm_1997} &
      \cite{schittkowski_nonlinear_1982, schittkowski_nonlinear_1982-1, schittkowski_convergence_1983, kraft_software_1988} &
      \cite{nash_newton-type_1984} & \cite{nocedal_numerical_2006}  & 
      \cite{powell_new_1970, nocedal_numerical_2006} &  \cite{steihaug_conjugate_1983, nocedal_numerical_2006} &
      \cite{conn_trust_2000, more_computing_1983} & \cite{gould_solving_1999, lenders_trlib:_2016} \\
    \hline
  \end{tabular}
  \label{tab:minimize-si}
\end{table}

>>>>

% PhD Mayo Clinic, started faculty position at BYU in 2001
By the late 1990s, discussions appeared on Matrix-SIG
expressing a desire for a complete scientific data analysis environment\cite{Travis-Keynote-2010}.
Travis Oliphant, a PhD student at the Mayo Clinic,
released a number of packages\cite{Travis-some-modules,Travis-enhance} 
that built on top of the Numeric array
package, and provided algorithms for signal processing, special
functions, sparse matrices, quadrature, optimization, Fast Fourier
Transforms, and more.  One of these packages, Multipack, was a set of
extension modules that wrapped Fortran and C libraries such as
ODEPACK, QUADPACK, and MINPACK, to solve and minimize nonlinear
equations, integrate differential equations, and fit splines (the
latter implemented by Pearu Peterson).  Robert Kern, then an
undergraduate student (and currently a SciPy core developer), provided
compilation instructions under Windows.
% 2001 PhD Tallinn Technical University
Around the same time, Pearu Peterson, a PhD student from Estonia,
released f2py\cite{Pearu-f2py-mail}, a command line tool for binding Python and Fortran
codes, and wrote modules for linear algebra and interpolation.
% Graduated 1999, Duke
Eric Jones, while a graduate student at Duke, wrote a number of
packages to support his dissertation, including a parallel job
scheduler and genetic optimizer.
% Graduated 1998, PhD from Brown, 97–99 postdoc at Harvard Medical
% School; at Harvard ever since
%
% https://mail.python.org/pipermail/scipy-dev/2001-November/000088.html
%
Gary Strangman, a postdoctoral fellow at Harvard Medical School,
published several descriptive and inferential statistical 
routines\cite{Strangman-modules}.

@@

% separate Figure numbering for SI
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

@@
